Qualifying Report

### 2019-01-20 Source/QualifyingReport/QualifyingReport.tex
QR edits 17-1-19

This section looks at my research aims and objectives given the literature review presented above. This section will then go on to explore any challenges to my research. I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives. As part of this I will identify target conferences where I could present my work. Finally, I look at potential risks and ethical concerns that may occur as part of my research.

@@

The aim of my research is to ultimately find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}. It is this area which is the focus of my research.

@@

I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia. Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language. I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is helpful to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments. Therefore we would like to see whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.

@@

\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline

@@

We compiled a series of tests which look a number of cognitive functions which have previously demonstrated a level of effectiveness at discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test). In addition to these tests, we have incorporated two language tasks in which a participants ability to generate spontaneous speech that we can then use to further analyse. As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.

@@

The aim of this task is to replicate and extend the work of Fraser and Orimaye who both looked at the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorise people into dementia or healthy categories using the samples given. We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors. Both datasets are longitudinal and have been shown to track language decline over time in different ways.

@@

Whilst we have divided participants into experimental groups we aim to investigate each participant individually. Because there is a large amount of heterogeneity in this population, we will initially be analysing each participants results in isolation with the participant acting as their own control. However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.

@@

The final output of this task will be a paper discussing the results of our longitudinal study from an language analysis perspective. The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.

>>>>

This section looks at my research aims and objectives given the literature review presented above. This section will then go on to explore any challenges to my research. I then present the work I have planned for the next two years, divided into distinct tasks with their own aims and objectives. As part of this I will identify target conferences where I could present my work. Finally, I look at potential risks and ethical concerns that could conceivably occur.

@@

The aim of my research is to ultimately to identify less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}. This will be the focus of my research.

@@

I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia. Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language. I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is helpful to establish that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments. Therefore we would like to clarify whether natural language generated outside the framework of a deliberate language generating task, has the same ability to discriminate.

@@

\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline

@@

We compiled a series of tests which look at a number of cognitive functions which have previously demonstrated a level of effectiveness in discriminating an MCI population from a healthy population (see literature review for a comprehensive list, explanation and rationale for each test). In addition to these tests, we have incorporated two language tasks in which a participant's ability to generate spontaneous speech that we can use to further analyse. As a planned future study is longitudinal, it is important that this battery of tests is repeatable but does not have significant practice effects that may interfere with our ability to track cognitive decline.

@@

The aim of this task is to replicate and extend the work of Fraser and Orimaye who both used the DementiaBank dataset and set out to use traditional machine learning techniques to see whether they could categorize people into dementia or healthy categories using the samples given. We also aim to replicate and extend the work of Le et al who used electronic versions of the novels of three authors. Both datasets are longitudinal and have been shown to track language decline over time in different ways.

@@

Whilst we have divided participants into experimental groups we aim to investigate each participant individually. Because there is a large amount of heterogeneity in this population, we will initially be analysing each participant's results in isolation with the participant acting as their own control. However, we will also be exploring the use of traditional machine learning and/or deep learning across cohorts to investigate whether each cohort has their own distinct 'language profile'.

@@

The final output of this task will be a paper discussing the results of our longitudinal study from a language analysis perspective. The target conference for this work will be the UK Alzheimer's Research conference in March 2021 with an expected submission date of December 2020, the Alzheimer's Association International Conference in July 2021 with an expected submission date of January 2021 and the Association of Computational Linguistics Conference (ACL) in July 2021 with an expected submission date of March 2021.
